---
title: "Probabilistic Analysis of Future Climate Change"
author: "DPM"
date: "2023-07-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(hector)
library(matilda)
library(dplyr)
library(tidyr)
library(ggplot2)
theme_set(theme_bw())
```

## Introduction

- Climate change presents an urgent global challenge, with the equilibrium climate sensitivity (ECS, the global temperature increase for a doubling of atmospheric CO2 concentrations) providing a key measure of the Earth's susceptibility to human influence.
- Estimates of ECS have relied on multiple lines of evidence, but significant uncertainty persists regarding the likely range.
- This work explores how omitting or incorporating different lines of ECS evidence impacts probabilistic projections from a simple climate model.

## Basic Hector Run

### Initialize Core

```{r configure-model}
ini_file <- system.file("input/hector_ssp245.ini", package = "hector")
core <- newcore(ini_file)
```

### Run Core

```{r run-core}
run(core)
results <- fetchvars(core, 2000:2300)
head(results)
```

### Plot of Basic Hector Output

```{r plot-hector-output}
ggplot(subset(results, variable == "CO2_concentration")) +
  aes(x = year, y = value) +
  geom_line() +
  ggtitle("CO2 Concentration")
```

## Matilda Run

### Generate parameter values

```{r generate-params}
set.seed(1)
param_values <- generate_params(core, draws = 10)
```

### Parameter histograms

```{r param-histograms}
# Reshape to long format
values_long <- param_values %>% 
  pivot_longer(
    cols = everything(), 
    names_to = "Column", 
    values_to = "Value"
  )

# Create a ggplot2 density plot
ggplot(values_long, aes(x = Value, fill = Column)) +
  geom_density(alpha = 0.6) +
  theme_minimal() +
  labs(x = "Value", y = "Density") +
  facet_wrap(~ Column, scales = "free", ncol = 1) +
  ggtitle("Density Curves for Each Column") +
  guides(fill = FALSE) +
  scale_y_continuous(n.breaks = 2)
```

### Run Hector model iteratively 

```{r run-hector, message=FALSE}
results <- iterate_hector(core = core,
                          params = param_values)
head(results)
```

### Plot model runs

```{r plot-model-runs}
ggplot(results, aes(x = year, y = value, group = run_number, color = as.factor(run_number))) +
  geom_line() +
  facet_wrap(~variable, scales = "free_y")
```

### Score model runs

```{r score-runs}
scores <- score_hruns(results, criterion_co2_obs(), score_ramp, w1 = 2, w2 = 20)

results_scored <- merge(results, scores, by = "run_number")
```

### Plot scored runs

```{r plot-scored-runs}
ggplot(data = results_scored) +
  geom_line(aes(x = year, y = value, group = run_number, color = weights)) +
  scale_color_continuous() +
  facet_wrap(~variable, scales = "free_y")
```

### Define and calculate metrics

```{r metrics}
my_metric <- new_metric(GLOBAL_TAS(), 2000:2100, mean)

metric_values <- metric_calc(results, my_metric)
```

### Weighted probabilistic projections

```{r projections, include=TRUE}
bins <- c(0, 1, 1.5, 2, 2.5, 3, 3.5, 4, Inf)

prob_calc(metric_values$metric_result, 
          bins = bins, 
          scores = scores$weights)
```

## Step 3
Talk about repeating the process for several unique ECS values




